### 2019-11-07 latex/vhcs.tex
Starting to write VHC section
Taking the lagrangian framework review from vnhcs and moving it to vhcs,
as vhcs come first in the thesis

\subsection{Lagrangian Framework Review}%
\label{sub:lagrangian_vhcs}

@@

\begin{defn}\label{defn:vhc_regular}
   A VHC is \textbf{regular} if system \textbf{TODO:
   reference to E-L mechanical system} with output \(e = h(q)\) has vector
   relative degree \(\{2,2,\ldots,2\}\).
   Equivalently, a VHC is regular if 
   \[
      dh_q D^{-1}(q)B(q)
   \]
   is nonsingular everywhere on \(\Gamma\).
\end{defn}

>>>>

Take an underactuated mechanical system with Lagrangian
\begin{equation}\label{eqn:lagrangian}
\mathcal{L}(q,\dot{q}) = \frac{1}{2} \dot{q}^T D(q) \dot{q} - P(q)
\end{equation}
where \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) is the configuration vector, 
the inertia matrix \(D(q) = D^T(q)\) is positive definite for all 
\(q \in \mathcal{Q}\), and \(P(q)\) is the potential of the system.

The equations of motion for this system can be computed from 
the Euler-Lagrange equations (\ref{eqn:el_eqns})
\begin{equation}\label{eqn:el_eqns}
   \frac{d}{dt} \left( \frac{\partial \mathcal{L}}{\partial \dot{q}}\right) -
   \frac{\partial \mathcal{L}}{\partial q} = B(q)\tau
\end{equation}
which result in (\ref{eqn:eom_lagrangian}), where
\(C : T\mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\) represents the
generalized coriolis and centrifugal forces,
\(B : \mathcal{Q} \rightarrow \mathbb{R}^{n \times m}\) is a 
\(C^1\) full rank control matrix, and \(\tau \in \mathbb{R}^{m}\) is the control 
input. \textbf{TODO: citation for E-L equations}

\begin{equation}\label{eqn:eom_lagrangian}
D(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla P(q) = B(q)\tau
\end{equation}

Suppose now that one wishes to enforce, through the control input \(\tau\), a
constraint on the generalized coordinates which depends only on \(q\) and not on
\(\dot{q}\). This type of constraint is called ``holonomic"; since the
constraint is enforced by control (as opposed to being enforced by the mechanics
of the system itself), we call the constraint ``virtual". This intuition leads
to the following mathematical definition as originally observed in 
\textbf{TODO: cite VHC definition paper}:

@@

\begin{defn}\label{defn:vhc_regular}
   A VHC is \textbf{regular} if system \textbf{TODO:
   reference to E-L mechanical system} with output \(e = h(q)\) has vector
   relative degree \(\{2,2,\ldots,2\}\).
   Equivalently, a VHC is regular if 
   \[
      dh_q D^{-1}(q)B(q)
   \]
   is nonsingular everywhere on \(\Gamma\).
\end{defn}
\textbf{TODO: cite VHC papers}

### 2019-11-07 latex/vnhcs.tex
Starting to write VHC section
Taking the lagrangian framework review from vnhcs and moving it to vhcs,
as vhcs come first in the thesis

%/----- Lagrange vs Hamilton -----/%
\subsection{A Hamiltonian Approach to Virtual Nonholonomic Constraints}
Take an underactuated mechanical system

\begin{equation}\label{eqn:eom_lagrangian}
M(q)\ddot{q} + C(q,\dot{q})\dot{q} + \nabla V(q) = B(q)\tau
\end{equation}
with Lagrangian
\begin{equation}\label{eqn:lagrangian}
\mathcal{L}(q,\dot{q}) = \frac{1}{2} \dot{q}^T M(q) \dot{q} - V(q)
\end{equation}
where \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) is the configuration vector, 
the inertia matrix \(M(q) = M^T(q)\) is positive definite for all 
\(q \in \mathcal{Q}\), \(V(q)\) is the potential of the system, 
\(B : \mathcal{Q} \rightarrow \mathbb{R}^{n \times m}\) is a 
\(C^1\) full rank control matrix, and \(\tau \in \mathbb{R}^{m}\) is the control 
input.

One can compute the Hamiltonian of this system by performing a Legendre transform on the Lagrangian. Define the conjugate of momenta for \(q\) by 
\[
p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
\]
Then, the Legendre transform is performed by taking
\[
\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
\]
It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is 
given by the total mechanical energy (\ref{eqn:hamiltonian}).
\begin{equation}\label{eqn:hamiltonian}
\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
\end{equation}

>>>>

%/----- Lagrange vs Hamilton -----/%
\subsection{A Hamiltonian Approach to Virtual Nonholonomic Constraints}
One can compute the Hamiltonian of the Lagrangian (\ref{eqn:lagrangian}) by 
performing a Legendre transform on the Lagrangian. 
First, define the conjugate of momenta for \(q\) by 
\[
p = \frac{\partial \mathcal{L}}{\partial \dot{q}} \in \mathbb{R}^n
\]
Then, the Legendre transform is performed by taking
\[
\mathcal{H}(q,p) = p^T \dot{q} - \mathcal{L}(q,\dot{q})
\]
It is easy to verify that the Hamiltonian for system (\ref{eqn:lagrangian}) is 
given by the total mechanical energy (\ref{eqn:hamiltonian}).
\begin{equation}\label{eqn:hamiltonian}
\mathcal{H}(q,p) = E(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
\end{equation}
